A saturation model for impaired learning with enhanced plasticity: Supplementary material

### 2014-01-09 MHCpaper/MHCsupp.tex
minor edits
VOR parameter fitting

% -*- TeX -*- -*- UK -*-
% ----------------------------------------------------------------
\documentclass[12pt]{article}
%Preamble:
\usepackage{a4wide}
\input{sl_preamble.tex}
\hypersetup{colorlinks,linkcolor=blue,citecolor=teal,urlcolor=teal}
\input{sl_graphics_preamble.tex}
\graphicspath{{"Figures/"}}
%
\usepackage{array}
\newcounter{tablerow}
\newcommand{\tablerowautorefname}{row}
\newenvironment{tabularn}[1]{\setcounter{tablerow}{0}\begin{tabular}{#1>{\refstepcounter{tablerow}}l}}{\end{tabular}}
%\usepackage{etoolbox}
%\AtBeginEnvironment{tabular}{\setcounter{tablerow}{0}}
%
% ----------------------------------------------------------------
% New commands etc.
\input{sl_definitions.tex}
\input{sl_symbols.tex}
%matrices
\newcommand{\I}{\mathbf{I}}
%vec of ones
%equilibrium distribution
\newcommand{\pr}{\mathbf{p}}
\newcommand{\eq}{\pr^\infty}
%other symbols
\newcommand{\w}{\mathbf{w}}
\newcommand{\W}{\mathbf{W}}
\newcommand{\frg}{\W^\mathrm{F}}
\newcommand{\M}{\mathbf{M}}
\newcommand{\F}{\boldsymbol{\Phi}}
\newcommand{\pot}{^{\text{pot}}}
\newcommand{\dep}{^{\text{dep}}}
\newcommand{\potdep}{^{\text{pot/dep}}}
\newcommand{\norm}{_0}
\newcommand{\inc}{_{\text{inc}}}
\newcommand{\dec}{_{\text{dec}}}
\newcommand{\incdec}{_{\text{inc/dec}}}
\newcommand{\wt}{_{\text{WT}}}
%\newcommand{\ko}{_{\text{D$^\mathrm{b}$K$^\mathrm{b}$-/-}}}
\newcommand{\KO}{D$^\mathrm{b}$K$^{\mathrm{b}-/-}$}
\newcommand{\ko}{_{\text{MHC}}}
%\newcommand{\KO}{MHC}
\newcommand{\tpre}{t_{\text{pre}}}
\newcommand{\ttrain}{t_{\text{train}}}
\newcommand{\lmax}{_{\text{max}}}
\newcommand{\lmin}{_{\text{min}}}
% ----------------------------------------------------------------
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Title info:
\title{A saturation model for impaired learning with enhanced plasticity: Supplementary material}
%
%% Author List:
%%
%\author{An Author$^a$, Another Author$^b$ and Yet Another Author$^c$\\
%%
%\small{\emph{$^a$ Address 1}}\\
%\small{\emph{$^b$ Address 2}}\\
%\small{\emph{$^c$ Address 3}}\\
%}

@@

For the serial, multistate and two state models, we will use the same value for the transition probabilities, $q$, for potentiation and depression in the wild-type as well as potentiation in the \KO\ mice.
We will use a larger value for $q$ for depression in the \KO\ mice.

For the cascade model, we will use the same value for the parameter $x$ (which controls the decay of transition rates, see \cite{Fusi2005cascade}) for potentiation and depression in the wild-type as well as potentiation in the \KO\ mice.
We will use a larger value for $x$ for depression in the \KO\ mice.

@@

This model is parameterised by the range of values, $q\in[q\lmin,q\lmax]$, for potentiation and depression.
We will use the same values for potentiation and depression in the wild-type as well as potentiation in the \KO\ mice.
We will use larger values for depression in the \KO\ mice.
In the main text, we consider a version of this model that has resource depletion for depression only, so that potentiation transition probabilities are unaffected by the number of potentiated synapses.
This is done by setting $q\pot\lmax=q\pot\lmin$.
The values of these parameters are listed in \autoref{tab:params}.

@@

This model is parameterised by the values of $x$ chosen for potentiation and depression.
We will use a larger value for depression in the \KO\ mice.
The values used are listed in \autoref{tab:params}.

@@

We will assume $f\dep\wt=f\dep\ko$.
This is because the effects of the mutation are well localised to the Purkinje cells, so the activity of the parallel and climbing fibres should not change very much.
Therefore the rates of potentiation and depression should not change very much either.
This implies that the mean synaptic weight in equilibrium differs between the \KO\ mice and wild type, but that is also suggested by the data regarding basal levels of phospho-GluR2 at serine 880.
As the mutation produces no change in baseline performance, there must be a compensatory mechanism somewhere else.
We will model this compensation as a simple linear offset, as could be produced by another population of neurons/synapses whose effect cancels with these neurons/synapses.

For the most part, we set $f\dep\norm=\half$, $f\dep\inc=f\dep\norm+\Delta f$ and $f\dep\dec=f\dep\norm-\Delta f$, with $\Delta f>0$.
We use the same values for wild-type and \KO\ mice for the reasons discussed above.
We will mostly treat gain-increase and decrease symmetrically, but this is not necessary.
We could adjust $r$ to keep $r\pot$ unchanged, if so desired, but this would only change the overall timescale and would not affect any of the qualitative comparisons that we are concerned with here.
The values of these parameters are listed in \autoref{tab:params}.

We are also assuming that the relation between VOR gain and mean synaptic weight is the same for \KO\ mice ant wild-type, except for a linear offset to compensate for the difference in equilibrium weights mentioned above.
This ensures that the qualitative questions mentioned above (enhancement or impairment of learning) will not be affected.

@@

We will try to gain some analytic insight to some of these models by looking at the slope of the learning curve at the start of gain-increase training.
This is proportional to the net-flux from the states with strong synaptic weight to the weaker states, measured using the transition probabilities for gain-increase but the equilibrium distribution for either untrained or gain-decrease, assuming that pre-training lasts long enough to reach the equilibrium distribution for gain-decrease.
This should be multiplied by the difference in synaptic weights.
But, as this quantity is constant within a model and doesn't change when comparing wild-type and \KO\ mice or with/without pre-training, it can be ignored.

@@

\newcommand{\captiontext}[1]{ can be found in \autoref{tr:#1} of \autoref{tab:params}.
  Time is measured in units of $1/r$, the mean time between plasticity events.
  (\ref{fig:#1_learn}) Learning curves for wild-type and \KO\ mice with and without pre-training.
  (\ref{fig:#1_learnS}) Learning curves restricted to gain-increase training.
  (\ref{fig:#1_eq}) Equilibrium distributions without training or with gain-increase/decrease training for (\ref{fig:#1_eq_WT}) wild-type and (\ref{fig:#1_eq_KO}) \KO\ mice,
  with states of weak synaptic strength to the left and strong states to the right.
  (\ref{fig:#1_pr}) Evolution of probability distributions for (\ref{fig:#1_pr_WT_nopre},\ref{fig:#1_pr_WT_pre}) wild-type and  (\ref{fig:#1_pr_KO_nopre},\ref{fig:#1_pr_KO_pre}) \KO\ mice without (\ref{fig:#1_pr_WT_nopre},\ref{fig:#1_pr_KO_nopre}) and with (\ref{fig:#1_pr_WT_pre},\ref{fig:#1_pr_KO_pre}) pre-training,
  with states of weak synaptic strength at the top and strong states underneath. }

@@

Now, consider the \KO\ mice, for which we define $\beta=q\pot/q\dep<1$, and we set $q\pot=q$.
Without pre-training:
%
\begin{equation}\label{eq:multiKOnopre}
  \Phi = 2{\Delta f} q\,\frac{(1-\beta)\beta^{M/2-1}}{1-\beta^M}.
\end{equation}
%
This will be smaller than \eqref{eq:multiWTnopre} if $\beta<\beta^*(M)$, where $\beta^*(M)$ is defined as the value at which they are equal.
This function is plotted in \autoref{fig:multistate_star}\ref{fig:multistate_betastar}, where we can see that it approaches 1 rapidly as we increase $M$.

@@

Let us define $\Delta f^*(\beta,M)$ to be the value at which \eqref{eq:multiKOnopre} and \eqref{eq:multiKOpre} are equal.
Values of $\Delta f$ that are larger than this (stronger training) will correspond to slower learning after pre-training.
As we would like pre-training to slow down learning in the wild-type but speed it up in the \KO\ mice, it would seem that we require $\Delta f^*(1,M) < \Delta f < \Delta f^*(\beta,M)$ (remembering that the wild-type corresponds to $\beta=1$).
As we see in \autoref{fig:multistate_star}\ref{fig:multistate_deltafstar}, $\Delta f^*(\beta,M) > \Delta f^*(1,M)$, so this range always exists.
We can see an example of this in \autoref{fig:multistate_med}\ref{fig:multistate_med_learnS}.

However, this is only the initial, instantaneous rate of change.
Let's look at \autoref{fig:multistate_strong}\ref{fig:multistate_strong_learnS}, for which $\Delta f > \Delta f^*(\beta,M)$, so that pre-training initially slows down learning for both wild-type and \KO\ mice.
We see that the pre-trained curve can rapidly catch up with the un-pre-trained one, due to the mass of probability concentrated at the end of the chain (see \autoref{fig:multistate_strong}\ref{fig:multistate_strong_eq}, red trace) drifting to the centre.
This happens sooner for the \KO\ mice than the wild-type, due to the stronger depressing transitions.
This means that there is an intermediate range of time-scales over which pre-training does slow down learning in the wild-type but speed it up in the \KO\ mice, even outside the desired range of values for $\Delta f$.
As we see in \autoref{fig:multistate_med}\ref{fig:multistate_med_learnS} (black curves), even when we are in this range, the period of time for which pre-training impairs learning in the wild-type can be very small.

\begin{figure}
 \begin{center}
 \begin{myenuma}
  \item\aligntop{\includegraphics[width=7cm]{multistate_betastar.eps}}\label{fig:multistate_betastar}
  \item\aligntop{\includegraphics[width=7cm]{multistate_deltafstar.eps}}\label{fig:multistate_deltafstar}
 \end{myenuma}
 \end{center}
  \caption[The functions $\beta^*(M)$ and $\Delta f^*(\beta,M)$]{The functions (\ref{fig:multistate_betastar}) $\beta^*(M)$, which describes when the \KO\ mice have impaired learning, and (\ref{fig:multistate_deltafstar}) $\Delta f^*(\beta,M)$ for $M=10$, which describes when pre-training enhances learning.}\label{fig:multistate_star}
\end{figure}

In conclusion, if we choose $\beta<\beta^*(M)$, $\Delta f > \Delta f^*(1,M)$ and look at intermediate time-scales, we will see that the \KO\ mice learn slower than wild-type without pre-training, and that pre-training speeds up learning in the \KO\ mice but slows it down in the wild-type.
However, in this case pre-training proceeds much faster in the \KO\ mice than wild-type (see \autoref{fig:multistate_med}\ref{fig:multistate_med_learn} and \autoref{fig:multistate_strong}\ref{fig:multistate_strong_learn}, lower curves), which is \emph{not} seen in the experiment.

@@

We can see from these formulae that, if we want pre-training to impair learning in wild-type, but enhance it in the \KO\ mice, the important thing is to have \emph{strong pre-training}, not strong training.
Thus, removing the symmetry between gain-increase and gain-decrease will not help with the most unrealistic feature of this model -- the fact that pre-training proceeds much faster in the \KO\ mice than wild-type.

Finally, we note that, in the case of medium or strong pre-training, we see positive curvature in the learning curve after pre-training for a longer time in wild-type than in the \KO\ mice (see \autoref{fig:multistate_med}\ref{fig:multistate_med_learnS} and \autoref{fig:multistate_strong}\ref{fig:multistate_strong_learnS}, dashed curves).
This is due to the mass of probability at the right end of the chain (see \autoref{fig:multistate_med}\ref{fig:multistate_med_eq} and \autoref{fig:multistate_strong}\ref{fig:multistate_strong_eq}, red trace) that is unable to contribute to change in synaptic weight until it has drifted to the centre.
This will last for a shorter time in the \KO\ mice, due to the enhanced transitions.
This will not be a feature of any of the other models that we study here.
Now, the curvature can be changed by the nonlinear relation between synaptic weight and VOR gain, but both curves would be affected in the same way.
The difference in these two curvatures is therefore a good feature of this model.

@@

Now, consider the \KO\ mice, for which we define $\beta=q\pot/q\dep<1$, and we set $q\pot=q$.
Without pre-training:
%
\begin{equation}\label{eq:binKOnopre}
  \Phi = \frac{2{\Delta f}\, q}{1+\beta},
\end{equation}
%
which is always larger that \eqref{eq:binWTnopre}, unlike what is seen in the experiment.
This is also seen in \autoref{fig:binary}\ref{fig:binary_learn},\ref{fig:binary_learnS} (solid curves).
As discussed above (below \eqref{eq:multiKOnopre}), the larger value of $q\dep=q/\beta$ has two effects.
Smaller $\beta$ will increase the probability of depression, speeding up learning, but it will also decrease the probability of being ready for depression, slowing down learning.
For the serial model, we argued that the first effect goes like $1/\beta$, whereas the second goes like $\beta^{M/2}$ to leading order.
Here $M=2$, so the leading part of the two effects will cancel.
The subleading effects (the normalisation of the probabilities) are responsible for the faster learning in the \KO\ mice.

@@

First, we assumed that the wild type had symmetric potentiation and depression.
Relaxing this assumption would correspond modelling the wild-type like the \KO\ mice, but with a different value for $\beta$.
We will still have $\beta\ko<\beta\wt$, and looking at \eqref{eq:binKOnopre} tell us that we would still have faster learning in the \KO\ mice.

@@

The only thing we have to worry about is when the wild-type catches up with the \KO\ mice.
As the total change will be larger for the wild-type than \KO\ mice, it must overtake eventually.
If this happens early enough, the initial period, where the \KO\ mice learns faster than wild-type, would not be seen in the experiment.
However, the timescale for this crossover would be similar to the timescale of the exponentials.
In the data, it looks like that timescale is longer than the gaps between successive measurements, so we needn't worry about this.

@@

Now, consider the \KO\ mice, for which we define $\beta=q\pot/q\dep<1$, and we set $q\pot=q$.
Without pre-training:
%
\begin{equation}\label{eq:multiLinKOnopre}
  \Phi = 2{\Delta f}\, q\,\frac{1-\beta^{M-1}}{1-\beta^M}.
\end{equation}
%
This decreases monotonically in the interval $\beta\in[0,1]$, therefore this will always be greater than \eqref{eq:multiLinWTnopre}.
This means that the \KO\ mice will initially learn faster than the wild-type.
However, as the wild-type will eventually catch up, and this can happen very quickly, as seen in \autoref{fig:multistate_lin}\ref{fig:multistate_lin_learn},\ref{fig:multistate_lin_learnS} (solid curves).

With pre-training:
%
\begin{equation}\label{eq:multiLinKOpre}
\begin{aligned}
  \Phi &= 4{\Delta f}\, q \, \frac{(1-2\Delta f)^{M-1} - \beta^{M-1}(1+2\Delta f)^{M-1}}
          {(1-2\Delta f)^M - \beta^M(1+2\Delta f)^M} \\
       &= 4{\Delta f}\, q\,\frac{1-\beta^{M-1}}{1-\beta^M} + \CO(\Delta f)^2.
\end{aligned}
\end{equation}
%
The ratio of this to \eqref{eq:multiLinKOnopre} takes its minimum value at the upper end of the interval $\Delta f \in \brk{0,\half}$, where it takes the value
%
\begin{equation}\label{eq:multiLinprenopre}
  \frac{\Phi_{\text{w/ pre}}}{\Phi_{\text{w/o pre}}} = \frac{1-\beta^M}{\beta-\beta^M}
   \longrightarrow \frac{M}{M-1} \quad \text{as} \quad \beta\to1,
\end{equation}
%
which is always greater than 1.
Therefore, pre-training will always enhance learning, for both \KO\ mice and wild-type, as seen in \autoref{fig:multistate_lin}\ref{fig:multistate_lin_learnS}.

@@

If we compare gain-increase learning in the \KO\ mice to wild-type, there are two effects: the increased transition rates speed up learning, but the equilibrium distribution is shifted to the depressed side (compare \autoref{fig:pooled_deponly}\ref{fig:pooled_deponly_eq}: \ref{fig:pooled_deponly_eq_WT} and \ref{fig:pooled_deponly_eq_KO}, blue traces), where there are fewer synapses available for depression and resources are depleted.
When resource depletion is sufficiently severe, the second effect dominates and the \KO\ mice learn slower than wild-type (see \autoref{fig:pooled_deponly}\ref{fig:pooled_deponly_learn},\ref{fig:pooled_deponly_learnS}, solid curves), which matches what is seen in the experiment.

Gain-decrease pre-training lessens the second effect, and results in the \KO\ mice learning faster than wild-type (see \autoref{fig:pooled_deponly}\ref{fig:pooled_deponly_learn},\ref{fig:pooled_deponly_learnS}), as seen in the experiment.

Gain-decrease pre-training will shift the distribution to the potentiated side, where there are more synapses available for depression and resources are more plentiful, for both the \KO\ mice and wild type (see \autoref{fig:pooled_deponly}\ref{fig:pooled_deponly_eq}, red traces).
This means that the pre-trained animals will learn faster then the untrained one for both \KO\ mice and wild-type (see \autoref{fig:pooled_deponly}\ref{fig:pooled_deponly_learnS}).
This differs from what is seen experimentally, where the pre-trained wild-type learns slower than the untrained one.

@@

In both of these, we see that the \KO\ mice are slower than wild-type without pre-training but faster with it.
This seems to be due to the fact that, without pre-training, very few synapses will be available for depression as most of them are already depressed (\autoref{fig:cascade_short}\ref{fig:cascade_short_eq}\ref{fig:cascade_short_eq_KO} and \ref{fig:cascade_short_pr}\ref{fig:cascade_short_pr_KO_nopre}).
With pre-training, some of them will now be potentiated (\autoref{fig:cascade_short}\ref{fig:cascade_short_eq}\ref{fig:cascade_short_eq_KO} and \ref{fig:cascade_short_pr}\ref{fig:cascade_short_pr_KO_pre}), and the enhanced depression can speed up learning.

>>>>

% -*- TeX -*- -*- UK -*-
% ----------------------------------------------------------------
\documentclass[12pt]{article}
%Preamble:
\usepackage{a4wide}
\input{sl_preamble.tex}
\hypersetup{colorlinks,linkcolor=blue,citecolor=teal,urlcolor=teal}
\input{sl_graphics_preamble.tex}
\graphicspath{{"Figures/"}}
%
\usepackage{array}
\newcounter{tablerow}
\newcommand{\tablerowautorefname}{row}
\newenvironment{tabularn}[1]{\setcounter{tablerow}{0}\begin{tabular}{#1>{\refstepcounter{tablerow}}l}}{\end{tabular}}
%\usepackage{etoolbox}
%\AtBeginEnvironment{tabular}{\setcounter{tablerow}{0}}
%
% ----------------------------------------------------------------
% New commands etc.
\input{sl_definitions.tex}
\input{sl_symbols.tex}
%matrices
\newcommand{\I}{\mathbf{I}}
%vec of ones
%equilibrium distribution
\newcommand{\pr}{\mathbf{p}}
\newcommand{\eq}{\pr^\infty}
%other symbols
\newcommand{\w}{\mathbf{w}}
\newcommand{\W}{\mathbf{W}}
\newcommand{\frg}{\W^\mathrm{F}}
\newcommand{\M}{\mathbf{M}}
\newcommand{\F}{\boldsymbol{\Phi}}
\newcommand{\pot}{^{\text{pot}}}
\newcommand{\dep}{^{\text{dep}}}
\newcommand{\potdep}{^{\text{pot/dep}}}
\newcommand{\norm}{_0}
\newcommand{\inc}{_{\text{inc}}}
\newcommand{\dec}{_{\text{dec}}}
\newcommand{\incdec}{_{\text{inc/dec}}}
\newcommand{\wt}{_{\text{WT}}}
%\newcommand{\ko}{_{\text{D$^\mathrm{b}$K$^\mathrm{b}$-/-}}}
\newcommand{\KO}{D$^\mathrm{b}$K$^{\mathrm{b}-/-}$}
\newcommand{\ko}{_{\text{MHC}}}
%\newcommand{\KO}{MHC}
\newcommand{\tpre}{t_{\text{pre}}}
\newcommand{\ttrain}{t_{\text{train}}}
\newcommand{\lmax}{_{\text{max}}}
\newcommand{\lmin}{_{\text{min}}}
% ----------------------------------------------------------------
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Title info:
\title{A saturation model for impaired learning with enhanced plasticity: Supplementary material}
%
%% Author List:
%%
%\author{An Author$^a$, Another Author$^b$ and Yet Another Author$^c$\\
%%
%\small{\emph{$^a$ Address 1}}\\
%\small{\emph{$^b$ Address 2}}\\
%\small{\emph{$^c$ Address 3}}\\
%}
\date{}

@@

For the serial, multistate and two state models, we will use the same value for the transition probabilities, $q$, for potentiation and depression in the wild-type as well as potentiation in the \KO\ models.
We will use a larger value for $q$ for depression in the \KO\ models.

For the cascade model, we will use the same value for the parameter $x$ (which controls the decay of transition rates, see \cite{Fusi2005cascade}) for potentiation and depression in the wild-type as well as potentiation in the \KO\ models.
We will use a larger value for $x$ for depression in the \KO\ models.

@@

This model is parameterised by the range of values, $q\in[q\lmin,q\lmax]$, for potentiation and depression.
We will use the same values for potentiation and depression in the wild-type as well as potentiation in the \KO\ models.
We will use larger values for depression in the \KO\ models.
In the main text, we consider a version of this model that has resource depletion for depression only, so that potentiation transition probabilities are unaffected by the number of potentiated synapses.
This is done by setting $q\pot\lmax=q\pot\lmin$.
The values of these parameters are listed in \autoref{tab:params}.

@@

This model is parameterised by the values of $x$ chosen for potentiation and depression.
We will use a larger value for depression in the \KO\ models.
The values used are listed in \autoref{tab:params}.

@@

We will assume $f\dep\wt=f\dep\ko$.
This is because the effects of the mutation are well localised to the Purkinje cells, so the activity of the parallel and climbing fibres should not change very much.
Therefore the rates of potentiation and depression should not change very much either.
This implies that the mean synaptic weight in equilibrium differs between the \KO\ models and wild type, but that is also suggested by the data regarding basal levels of phospho-GluR2 at serine 880.
As the mutation produces no change in baseline performance, there must be a compensatory mechanism somewhere else.
We will model this compensation as a simple linear offset, as could be produced by another population of neurons/synapses whose effect cancels with these neurons/synapses.

For the most part, we set $f\dep\norm=\half$, $f\dep\inc=f\dep\norm+\Delta f$ and $f\dep\dec=f\dep\norm-\Delta f$, with $\Delta f>0$.
We use the same values for wild-type and \KO\ models for the reasons discussed above.
We will mostly treat gain-increase and decrease symmetrically, but this is not necessary.
We could adjust $r$ to keep $r\pot$ unchanged, if so desired, but this would only change the overall timescale and would not affect any of the qualitative comparisons that we are concerned with here.
The values of these parameters are listed in \autoref{tab:params}.

We are also assuming that the relation between VOR gain and mean synaptic weight is the same for \KO\ mice and wild-type, except for a linear offset to compensate for the difference in equilibrium weights mentioned above.
This ensures that the qualitative questions mentioned above (enhancement or impairment of learning) will not be affected.

@@

We will try to gain some analytic insight to some of these models by looking at the slope of the learning curve at the start of gain-increase training.
This is proportional to the net-flux from the states with strong synaptic weight to the weaker states, measured using the transition probabilities for gain-increase but the equilibrium distribution for either untrained or gain-decrease, assuming that pre-training lasts long enough to reach the equilibrium distribution for gain-decrease.
This should be multiplied by the difference in synaptic weights.
But, as this quantity is constant within a model and doesn't change when comparing wild-type and \KO\ models or with/without pre-training, it can be ignored.

@@

\newcommand{\captiontext}[1]{ can be found in \autoref{tr:#1} of \autoref{tab:params}.
  Time is measured in units of $1/r$, the mean time between plasticity events.
  (\ref{fig:#1_learn}) Learning curves for wild-type and \KO\ models with and without pre-training.
  (\ref{fig:#1_learnS}) Learning curves restricted to gain-increase training.
  (\ref{fig:#1_eq}) Equilibrium distributions without training or with gain-increase/decrease training for (\ref{fig:#1_eq_WT}) wild-type and (\ref{fig:#1_eq_KO}) \KO\ models,
  with states of weak synaptic strength to the left and strong states to the right.
  (\ref{fig:#1_pr}) Evolution of probability distributions for (\ref{fig:#1_pr_WT_nopre},\ref{fig:#1_pr_WT_pre}) wild-type and  (\ref{fig:#1_pr_KO_nopre},\ref{fig:#1_pr_KO_pre}) \KO\ models without (\ref{fig:#1_pr_WT_nopre},\ref{fig:#1_pr_KO_nopre}) and with (\ref{fig:#1_pr_WT_pre},\ref{fig:#1_pr_KO_pre}) pre-training,
  with states of weak synaptic strength at the top and strong states underneath. }

@@

Now, consider the \KO\ models, for which we define $\beta=q\pot/q\dep<1$, and we set $q\pot=q$.
Without pre-training:
%
\begin{equation}\label{eq:multiKOnopre}
  \Phi = 2{\Delta f} q\,\frac{(1-\beta)\beta^{M/2-1}}{1-\beta^M}.
\end{equation}
%
This will be smaller than \eqref{eq:multiWTnopre} if $\beta<\beta^*(M)$, where $\beta^*(M)$ is defined as the value at which they are equal.
This function is plotted in \autoref{fig:multistate_star}\ref{fig:multistate_betastar}, where we can see that it approaches 1 rapidly as we increase $M$.

@@

Let us define $\Delta f^*(\beta,M)$ to be the value at which \eqref{eq:multiKOnopre} and \eqref{eq:multiKOpre} are equal.
Values of $\Delta f$ that are larger than this (stronger training) will correspond to slower learning after pre-training.
As we would like pre-training to slow down learning in the wild-type but speed it up in the \KO\ models, it would seem that we require $\Delta f^*(1,M) < \Delta f < \Delta f^*(\beta,M)$ (remembering that the wild-type corresponds to $\beta=1$).
As we see in \autoref{fig:multistate_star}\ref{fig:multistate_deltafstar}, $\Delta f^*(\beta,M) > \Delta f^*(1,M)$, so this range always exists.
We can see an example of this in \autoref{fig:multistate_med}\ref{fig:multistate_med_learnS}.

However, this is only the initial, instantaneous rate of change.
Let's look at \autoref{fig:multistate_strong}\ref{fig:multistate_strong_learnS}, for which $\Delta f > \Delta f^*(\beta,M)$, so that pre-training initially slows down learning for both wild-type and \KO\ models.
We see that the pre-trained curve can rapidly catch up with the un-pre-trained one, due to the mass of probability concentrated at the end of the chain (see \autoref{fig:multistate_strong}\ref{fig:multistate_strong_eq}, red trace) drifting to the centre.
This happens sooner for the \KO\ models than the wild-type, due to the stronger depressing transitions.
This means that there is an intermediate range of time-scales over which pre-training does slow down learning in the wild-type but speed it up in the \KO\ models, even outside the desired range of values for $\Delta f$.
As we see in \autoref{fig:multistate_med}\ref{fig:multistate_med_learnS} (black curves), even when we are in this range, the period of time for which pre-training impairs learning in the wild-type can be very small.

\begin{figure}
 \begin{center}
 \begin{myenuma}
  \item\aligntop{\includegraphics[width=7cm]{multistate_betastar.eps}}\label{fig:multistate_betastar}
  \item\aligntop{\includegraphics[width=7cm]{multistate_deltafstar.eps}}\label{fig:multistate_deltafstar}
 \end{myenuma}
 \end{center}
  \caption[The functions $\beta^*(M)$ and $\Delta f^*(\beta,M)$]{The functions (\ref{fig:multistate_betastar}) $\beta^*(M)$, which describes when the \KO\ models have impaired learning, and (\ref{fig:multistate_deltafstar}) $\Delta f^*(\beta,M)$ for $M=10$, which describes when pre-training enhances learning.}\label{fig:multistate_star}
\end{figure}

In conclusion, if we choose $\beta<\beta^*(M)$, $\Delta f > \Delta f^*(1,M)$ and look at intermediate time-scales, we will see that the \KO\ models learn slower than wild-type without pre-training, and that pre-training speeds up learning in the \KO\ models but slows it down in the wild-type.
However, in this case pre-training proceeds much faster in the \KO\ models than wild-type (see \autoref{fig:multistate_med}\ref{fig:multistate_med_learn} and \autoref{fig:multistate_strong}\ref{fig:multistate_strong_learn}, lower curves), which is \emph{not} seen in the experiment.

@@

We can see from these formulae that, if we want pre-training to impair learning in wild-type, but enhance it in the \KO\ models, the important thing is to have \emph{strong pre-training}, not strong training.
Thus, removing the symmetry between gain-increase and gain-decrease will not help with the most unrealistic feature of this model -- the fact that pre-training proceeds much faster in the \KO\ models than wild-type.

Finally, we note that, in the case of medium or strong pre-training, we see positive curvature in the learning curve after pre-training for a longer time in wild-type than in the \KO\ models (see \autoref{fig:multistate_med}\ref{fig:multistate_med_learnS} and \autoref{fig:multistate_strong}\ref{fig:multistate_strong_learnS}, dashed curves).
This is due to the mass of probability at the right end of the chain (see \autoref{fig:multistate_med}\ref{fig:multistate_med_eq} and \autoref{fig:multistate_strong}\ref{fig:multistate_strong_eq}, red trace) that is unable to contribute to change in synaptic weight until it has drifted to the centre.
This will last for a shorter time in the \KO\ models, due to the enhanced transitions.
This will not be a feature of any of the other models that we study here.
Now, the curvature can be changed by the nonlinear relation between synaptic weight and VOR gain, but both curves would be affected in the same way.
The difference in these two curvatures is therefore a good feature of this model.

@@

Now, consider the \KO\ models, for which we define $\beta=q\pot/q\dep<1$, and we set $q\pot=q$.
Without pre-training:
%
\begin{equation}\label{eq:binKOnopre}
  \Phi = \frac{2{\Delta f}\, q}{1+\beta},
\end{equation}
%
which is always larger that \eqref{eq:binWTnopre}, unlike what is seen in the experiment.
This is also seen in \autoref{fig:binary}\ref{fig:binary_learn},\ref{fig:binary_learnS} (solid curves).
As discussed above (below \eqref{eq:multiKOnopre}), the larger value of $q\dep=q/\beta$ has two effects.
Smaller $\beta$ will increase the probability of depression, speeding up learning, but it will also decrease the probability of being ready for depression, slowing down learning.
For the serial model, we argued that the first effect goes like $1/\beta$, whereas the second goes like $\beta^{M/2}$ to leading order.
Here $M=2$, so the leading part of the two effects will cancel.
The subleading effects (the normalisation of the probabilities) are responsible for the faster learning in the \KO\ models.

@@

First, we assumed that the wild type had symmetric potentiation and depression.
Relaxing this assumption would correspond modelling the wild-type like the \KO\ mice, but with a different value for $\beta$.
We will still have $\beta\ko<\beta\wt$, and looking at \eqref{eq:binKOnopre} tell us that we would still have faster learning in the \KO\ models.

@@

The only thing we have to worry about is when the wild-type catches up with the \KO\ models.
As the total change will be larger for the wild-type than \KO\ models, it must overtake eventually.
If this happens early enough, the initial period, where the \KO\ models learns faster than wild-type, would not be seen in the experiment.
However, the timescale for this crossover would be similar to the timescale of the exponentials.
In the data, it looks like that timescale is longer than the gaps between successive measurements, so we needn't worry about this.

@@

Now, consider the \KO\ models, for which we define $\beta=q\pot/q\dep<1$, and we set $q\pot=q$.
Without pre-training:
%
\begin{equation}\label{eq:multiLinKOnopre}
  \Phi = 2{\Delta f}\, q\,\frac{1-\beta^{M-1}}{1-\beta^M}.
\end{equation}
%
This decreases monotonically in the interval $\beta\in[0,1]$, therefore this will always be greater than \eqref{eq:multiLinWTnopre}.
This means that the \KO\ models will initially learn faster than the wild-type.
However, as the wild-type will eventually catch up, and this can happen very quickly, as seen in \autoref{fig:multistate_lin}\ref{fig:multistate_lin_learn},\ref{fig:multistate_lin_learnS} (solid curves).

With pre-training:
%
\begin{equation}\label{eq:multiLinKOpre}
\begin{aligned}
  \Phi &= 4{\Delta f}\, q \, \frac{(1-2\Delta f)^{M-1} - \beta^{M-1}(1+2\Delta f)^{M-1}}
          {(1-2\Delta f)^M - \beta^M(1+2\Delta f)^M} \\
       &= 4{\Delta f}\, q\,\frac{1-\beta^{M-1}}{1-\beta^M} + \CO(\Delta f)^2.
\end{aligned}
\end{equation}
%
The ratio of this to \eqref{eq:multiLinKOnopre} takes its minimum value at the upper end of the interval $\Delta f \in \brk{0,\half}$, where it takes the value
%
\begin{equation}\label{eq:multiLinprenopre}
  \frac{\Phi_{\text{w/ pre}}}{\Phi_{\text{w/o pre}}} = \frac{1-\beta^M}{\beta-\beta^M}
   \longrightarrow \frac{M}{M-1} \quad \text{as} \quad \beta\to1,
\end{equation}
%
which is always greater than 1.
Therefore, pre-training will always enhance learning, for both \KO\ models and wild-type, as seen in \autoref{fig:multistate_lin}\ref{fig:multistate_lin_learnS}.

@@

If we compare gain-increase learning in the \KO\ models to wild-type, there are two effects: the increased transition rates speed up learning, but the equilibrium distribution is shifted to the depressed side (compare \autoref{fig:pooled_deponly}\ref{fig:pooled_deponly_eq}: \ref{fig:pooled_deponly_eq_WT} and \ref{fig:pooled_deponly_eq_KO}, blue traces), where there are fewer synapses available for depression and resources are depleted.
When resource depletion is sufficiently severe, the second effect dominates and the \KO\ models learn slower than wild-type (see \autoref{fig:pooled_deponly}\ref{fig:pooled_deponly_learn},\ref{fig:pooled_deponly_learnS}, solid curves), which matches what is seen in the experiment.

Gain-decrease pre-training lessens the second effect, and results in the \KO\ models learning faster than wild-type (see \autoref{fig:pooled_deponly}\ref{fig:pooled_deponly_learn},\ref{fig:pooled_deponly_learnS}), as seen in the experiment.

Gain-decrease pre-training will shift the distribution to the potentiated side, where there are more synapses available for depression and resources are more plentiful, for both the \KO\ models and wild type (see \autoref{fig:pooled_deponly}\ref{fig:pooled_deponly_eq}, red traces).
This means that the pre-trained animals will learn faster then the untrained one for both \KO\ models and wild-type (see \autoref{fig:pooled_deponly}\ref{fig:pooled_deponly_learnS}).
This differs from what is seen experimentally, where the pre-trained wild-type learns slower than the untrained one.

@@

In both of these, we see that the \KO\ models are slower than wild-type without pre-training but faster with it.
This seems to be due to the fact that, without pre-training, very few synapses will be available for depression as most of them are already depressed (\autoref{fig:cascade_short}\ref{fig:cascade_short_eq}\ref{fig:cascade_short_eq_KO} and \ref{fig:cascade_short_pr}\ref{fig:cascade_short_pr_KO_nopre}).
With pre-training, some of them will now be potentiated (\autoref{fig:cascade_short}\ref{fig:cascade_short_eq}\ref{fig:cascade_short_eq_KO} and \ref{fig:cascade_short_pr}\ref{fig:cascade_short_pr_KO_pre}), and the enhanced depression can speed up learning.
